---
title: "Index Swapping"
output:
  html_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd('/home/christian/gdrive/School/bi622/index_hopping/counts')
library(jsonlite)
library(reshape2)
library(ggpubr)
library(tidyr)
library(scales)
```

First lets load the JSON object from the python script and link some names to make life easier.

```{r, echo=TRUE}
base=fromJSON('/home/christian/gdrive/School/bi622/index_hopping/counts/count_0')
meta=base$metadata
codes=base$code_set
qual=base$qual_info
tags=base$tag_counts
rd_avg=t(qual$read_avg_dist)
```

Print the metadata for posterity.
```{r}
print(meta)
```

# Part 1 
## A. 
Turn in the 8 histograms.
```{r, include=FALSE}
rdf <- data.frame(rd_avg)
rdf_melt <- melt(rd_avg)

rr1 <- ggplot(data=subset(rdf_melt, Var2==1), aes(x=Var1, y=value))+
 geom_col(width = 1, color = "black")+
 scale_y_log10()+
 xlab("Avg Phred Per Read")+ylab("Counts")

ri1 <- ggplot(data=subset(rdf_melt, Var2==2), aes(x=Var1, y=value))+
 geom_col(width = 1, color = "black")+
 scale_y_log10()+
 xlab("Avg Phred Per Read")+ylab("Counts")

ri2 <- ggplot(data=subset(rdf_melt, Var2==3), aes(x=Var1, y=value))+
 geom_col( width = 1, color = "black")+
 scale_y_log10()+
 xlab("Avg Phred Per Read")+ylab("Counts")

rr2 <- ggplot(data=subset(rdf_melt, Var2==4), aes(x=Var1, y=value))+
 geom_col(width = 1, color = "black")+
 scale_y_log10()+
 xlab("Avg Phred Per Read")+ylab("Counts")

pos_melt=melt(qual$pos_avg_dist)
pos_melt=transform(pos_melt, i = ave(L1, L1, FUN = seq_along)) 


pr1 <- ggplot(data=subset(pos_melt, L1==1), aes(x=i, y=value))+
 geom_col(width = 1, color = "black")+
 xlab("Position")+ylab("Avg Base Phred")

pi1 <- ggplot(data=subset(pos_melt, L1==2), aes(x=i, y=value))+
 geom_col(width = 1, color = "black")+
 xlab("Position")+ylab("Avg Base Phred")

pi2 <- ggplot(data=subset(pos_melt, L1==3), aes(x=i, y=value))+
 geom_col(width = 1, color = "black")+
 xlab("Position")+ylab("Avg Base Phred")

pr2 <- ggplot(data=subset(pos_melt, L1==4), aes(x=i, y=value))+
 geom_col(width = 1, color = "black")+
 xlab("Position")+ylab("Avg Base Phred")
```
**Average Score per Base Position**
```{r, fig.width=12}
ggarrange(rr1, ri1, ri2, rr2 + rremove("x.text"), 
          labels = c("R1", "I1", "I2", "R2"),
          ncol = 2, nrow = 2)
```
**Average Score per Read**
```{r, fig.width=12}
ggarrange(pr1, pi1, pi2, pr2 + rremove("x.text"), 
          labels = c("R1", "I1", "I2", "R2"),
          ncol = 2, nrow = 2)

```

The average quality of reads was `r paste(c('R1:','I1:','I2:','R2:'), qual$run_avg)`

And the total number of reads was `r paste(c('R1:','I1:','I2:','R2:'),qual$rec_num)`

## B. 
What is a good quality score cutoff for index reads and pairs to utilize for sample 
identification and downstream analysis, respectively? 

**A Cutoff of 20 would seem to be good.**

## C. 
How many indexes have Undetermined (N) base calls? (Utilize your command line 
tool knowledge. Summit the command you used. CHALLENGE: use a one line command)
```{bash, eval=FALSE}
ssh hpc "zcat ~/jobs/622_index/1294*.gz | sed -n '2~4p' | awk 'BEGIN {Ns=0} /N/{Ns+=1} END {print Ns}'"
```

`13499075`

## D. 
What do the averaged Quality Scores across the reads tell you? Interpret your data 
specifically. 

**That the vast majority of reads fall above a score on 15. And follow a sigmoid curve.**


# Part 2. 
Write a program to document index swapping and number of reads retained per sample. 

Load the JSON for the good cutoff and link some names to make life easier.

```{r, echo=TRUE}
base=fromJSON('/home/christian/gdrive/School/bi622/index_hopping/counts/count_20')
meta=base$metadata
codes=base$code_set
qual=base$qual_info
tags=base$tag_counts
rd_avg=t(qual$read_avg_dist)
```

Print the metadata for posterity.

```{r}
print(meta)
```
## A. 
How many reads are retained for each expected index pair? What is the percentage? 

**Reads for each of the correct barcode pairs: "I1_I2 Count proportion"**

`r paste(codes$known, tags[codes$known],unlist(tags[codes$known])/qual$rec_num[1])`

**Reads for each of the correct barcode pairs: "I1_I2 Count proportion"**

`r paste(codes$known, tags[codes$known],unlist(tags[codes$known])/qual$rec_num[1])`

## B.
How many reads are indicative of index swapping? 

**The sum of swapped reads is `r sum(unlist(tags[codes$swap]))` or the proportion is  `r sum(unlist(tags[codes$swap]))/qual$rec_num[1]` of all reads**

## C.
Create a distribution of swapped indexes. What does this heatmap tell you/what is your interpretation of this data? 

```{r, fig.height=8, fig.width=10}
all_codes=c(codes$known,codes$swaped)
code_counts=tags[all_codes]
code_counts_melt=melt(code_counts)
code_counts_sep=separate(data = code_counts_melt, col = L1, into = c("I1", "I2"), sep = "\\_")
names(code_counts_sep)=c('counts', 'I1', 'I2')
resistor <- c("#000000", "#663300", "#FF0000", "#FF6c00", "#FFFF00", "#00FF00", "#0000FF", "#FF00FF","#7D7D7D","#FFFFFF")

swap <- ggplot(data=code_counts_sep, aes(y=I1, x=I2, fill=log10(counts)))+
geom_tile(stat="identity")+
scale_fill_gradientn(colours = resistor, limits=c(0, 10), breaks= 1:10)+
xlab("Index 2")+ylab("Index 1")+
theme(axis.text.x = element_text(angle = 90 ,hjust = 1))
swap
```

**It is interesting, there is less index swapping that I expected. However the index swappiing that does occur is clearly not random. The 2 indexes with the highest swapping are recipical to one another. And one library that had 10x more reads than any other library showed an increased swapping across the board.**
